---
title: "Coding Notes Iterations and Functions"
author: "Mamata K C"
date: "2025-03-27"
output: pdf_document
toc: true
---


```{r}
# Convert Fahrenheit to Celsius mathematically using the formula
(5*(32-32)/9)
(5*(40-32)/9)
(5*(80-32)/9)
(5*(120-32)/9)

```

- problem in reproducibility is copy paste might cause error in complex examples.
- to solve this problem, we can write a function for the formula.



# Functions
```{r}
# Writing a function to convert Fahrenheit to Celsius

F_to_C <- function(f_temp){
  celsius <- (5*(f_temp-32)/9)
  return(celsius)
}


F_to_C(32)
F_to_C(46)
F_to_C(90)
```

- In this function, we are creating a function such that we want to ask what the Fahrenheit temperature is and we want function to do the calculation and return celsius value.

```{r}
# Writing a function to convert Celsius to Fahrenheit 

C_To_F <- function(c_temp){
  fahrenheit <- (c_temp*(9/5)+32)
  return(fahrenheit)
}

C_To_F(2)
C_To_F(30)
```



# Iterations

```{r}
# iteration function in base R
rep("A",3)   #repeat A three times
rep(c("A","B"), 10) #repeat A and B ten times
rep(c(1,2,5,3),4, each = 5) #repeat each number in sequence of 4 five times

1:7
seq(from = 1, to = 7) #does same thing as 1:7
seq(from = 0, to = 10, by = 2) #can get into complicated examples such as maintaining gap within sequence

# combined seq() and rep()
rep(seq(from = 0, to = 10, by = 2), times = 3, each = 2)

LETTERS
seq_along(LETTERS)
```

# The for loop

```{r}
for (i in 1:10){
  print(i*2)      #takes each value from 1 to 10 and multiplies it with 2 and prints the value
}

for (i in -30:100){
  result <- F_to_C(i)
  print(result)
}
```
- The result we obtained are just displayed in console but they are not saved as any object. 

```{r}
celsius.df <- NULL   #create a null object
for (i in -30:100){
  result <- data.frame(F_to_C(i), i)    #create a dataframe named result with two columns (one as input, other as output of function)
  celsius.df <- rbind.data.frame(celsius.df, result)  #each time iteration happens, previous result is row bind to new result until iteration is completed
}

celsius.df
```

# Practical example 

```{r}
library(ggplot2)
library(drc)
library(tidyverse)

EC50.data <- read.csv("EC50_all.csv")


isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == "ILSO_5-41c"] ~ 
        EC50.data$conc[EC50.data$is == "ILSO_5-41c"], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
# outputs the summary of the paramters including the estimate, standard
# error, t-value, and p-value outputs it into a data frame called
# summary.mef.fit for 'summary of fit'
summary.fit <- data.frame(summary(isolate1)[[3]])
# outputs the summary of just the EC50 data including the estimate, standard
# error, upper and lower bounds of the 95% confidence intervals around the
# EC50
EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]


nm <- unique(EC50.data$is)

for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
  print(nm[[i]])
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    
    EC50
}   


#to solve the problem of results being displayed only in colsole

EC50.ll4 <- NULL
nm <- unique(EC50.data$is)
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
  print(nm[[i]])
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    isolate.ec_1 <- data.frame(nm[[i]], EC50)
    EC50.ll4 <- rbind.data.frame(EC50.ll4, isolate.ec_1)
    EC50
}



```

```{r}
#Another way to do the same thing using tidyverse

EC50.data %>%
  group_by(is) %>%
  nest() %>%    #nest allows you to make a sub dataframe within a dataframe
  mutate(ll.4.mod = map(data, ~drm(.$relgrowth ~ .$conc,     #map function tells us we want to iterate (it is basically a for loop contained within a function) (. symbol inherits all of the columns from data)
                              fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                         names = c("Slope", "Lower", "Upper", "EC50"))))) %>%
  mutate(ec50 = map(ll.4.mod, ~ED(., 
                              respLev = c(50), 
                              type = "relative",
                              interval = "delta")[[1]])) %>%
  unnest(ec50)

```

