---
title: "CodingChallenge7_LinearModels"
author: "Mamata K C"
date: "2025-04-03"
output:
  html_document:
    toc: true
    df_print: paged
  md_document:
    variant: gfm
  pdf_document:
    toc: true
---


Linear Model – 25 pts

PLEASE READ THIS BEFORE CONTINUING

This assignment will help you practice writing, executing, and interpreting a linear model in R. It will also involve some more practice with GitHub. You may collaborate with a partner to enhance your learning experience. Please ensure the following:

•	Collaboration: If you work with a partner, include both names on the final submission by editing the YAML header.
•	Submission: Only one person should submit the assignment to Canvas in a Word document or .pdf file generated through R markdown. Additionally, you should provide a link to your GitHub, where the assignment should be viewable by rendering it as a GitHub-flavored markdown file. 
•	Setup: It is also assumed you already have a GitHub repository for this class. 
•	Time: This should take you no longer than the class period to complete. 


# 1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
STAND$Treatment <- as.factor(STAND$Treatment) # example shown here. 


```{r}
#loading all the necessary libraries

library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
library(stringr)

datum <- read.csv("PlantEmergence.csv") #loading data using relative file path
head(datum)

datum$Treatment <- as.factor(datum$Treatment) #setting treatment as factor
datum$DaysAfterPlanting <- as.factor(datum$DaysAfterPlanting) #setting DaysAfterPlanting as factor
datum$Rep <- as.factor(datum$Rep) #setting Rep as factor

```

# 2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 


```{r}

lm1 <- lm(Emergence~Treatment+DaysAfterPlanting+Treatment:DaysAfterPlanting, data = datum) #running linear model
summary(lm1)
anova(lm1) #anova

```


# 3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

No, we don't need to fit the interaction term in the linear model because interaction was found to be non significant.

```{r}
lm2 <- lm(Emergence~Treatment+DaysAfterPlanting, data = datum)
summary(lm2)
confint(lm2)
anova(lm2)
```


Interpretation: 
Intercept: Estimated emergence for Treatment 1 at day 7 is 182.163.
Coeficient for Treatment 2: We found that Treatment 2 has 134.531 (+-6.77;+-95%C.I.) plants lesser emergence than the Treatment 1 (p value < 2e-16).


# 4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.


```{r}
lsmeans <- emmeans(lm2, ~Treatment) 
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE)
Results_lsmeans
```

Interpretation: least squared means are the means estimated by linear model. According the result, Treatment 2 is significantly different from all other Treatments. Treatment 7 is significantly different from Treamtents 3, 5 and 6. All other remaining treatments 1, 3, 4, 5, 6, 8, 9 are not significantly different from each other.


# 5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

plot_cldbars_onefactor(lm2,"Treatment")

```



Significance of letters: Letter c denotes Treatment 2 is significantly different from all other Treatments as they have letters a or ab. Treatment 7 has letter b which means it is significantly different from Treamtents 3, 5 and 6 as they have letters a. However, it is not significantly different from Treatments 1, 4, 8, 9 as they have letter ab. All other remaining treatments 1, 3, 4, 5, 6, 8, 9 are not significantly different from each other.


# 6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.


[Link to my GitHub](https://github.com/mamata2423/Reproducibility_Assignments.git)