---
title: "CodingNotesLinearModels"
author: "Mamata K C"
date: "2025-04-03"
output: pdf_document
---


```{r}
#loading all necessary packages
library(ggplot2)
library(tidyverse)
library(multcomp)
library(lme4)
library(emmeans)

data("mtcars") #loading data
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = lm, se = FALSE, color = "grey") +
  geom_point(aes(color = wt)) +
  xlab("Weight") + 
  ylab("Miles per gallon") +
  scale_colour_gradient(low = "forestgreen", high = "black") +
  theme_classic()

```

- to know if there is relationship between miles per gallon and weight, we run linear model which gives the estimate for our intercept.

```{r}
lm1 <- lm(mpg~wt, data = mtcars) #running linear model
summary(lm1)

anova(lm(mpg~wt, data = mtcars)) #running an anova (gives anova table)

cor.test(mtcars$wt, mtcars$mpg) #correlation test
```
- The value of multiple R squared gives the idea that 75% of the variation in miles per gallon is due to weight.
- linear model, ANOVA or correlation, all are basically the same thing, gives same p value and all are based on regression.


```{r}
bull.rich <- read.csv("Bull_richness.csv") #loading data
bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.") %>%
  ggplot(aes(x = Fungicide, y = richness)) + 
  geom_boxplot()
```

- Again, looking at the boxplot, it looks like there is a difference between the treatments so we can do t.test (ANOVA with two groups).

```{r}
bull.rich.sub <- bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.") #subset 1

t.test(richness~Fungicide, data = bull.rich.sub) #t test

summary(lm(richness~Fungicide, data = bull.rich.sub)) #linear model

anova(lm(richness~Fungicide, data = bull.rich.sub))
```

- Again, all of the tests give same result.

```{r}
bull.rich.sub2 <- bull.rich %>%
  filter(Fungicide == "C" & Treatment == "Conv." & Crop == "Corn") #subset 2

bull.rich.sub2$GrowthStage <- factor(bull.rich.sub2$GrowthStage, levels = c("V6", "V8", "V15"))

ggplot(bull.rich.sub2, aes(x = GrowthStage, y = richness)) +
  geom_boxplot() #plotting the subset data

lm.growth <- lm(richness ~ GrowthStage, data = bull.rich.sub2) #linear model
summary(lm.growth)

#ANOVAs
anova(lm.growth)

summary(aov(richness ~ GrowthStage, data = bull.rich.sub2))

```

- All the above resutls have same p value again.
- To find out which groups are different from each other, we can do post hoc test.



```{r}
lsmeans <- emmeans(lm.growth, ~GrowthStage) 
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE)
Results_lsmeans
```
- emmeans gives means of our model distribution for each categories and their confidence interval.
- also outputs which group are different from each other.



```{r}
bull.rich.sub3 <- bull.rich %>%
  filter(Treatment == "Conv." & Crop == "Corn")  #subset3

bull.rich.sub3$GrowthStage <- factor(bull.rich.sub3$GrowthStage, levels = c("V6", "V8", "V15"))

lm.interactions <- lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3) #interaction between growthstage and fungicide
summary(lm.interactions)
anova(lm.interactions)


lsmeans <- emmeans(lm.interactions, ~Fungicide|GrowthStage) # effect of fungicide within growthstage
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) 
Results_lsmeans

ggplot(bull.rich.sub3, aes(x = GrowthStage, y = richness, color = Fungicide)) + 
  geom_boxplot()
```

```{r}
lme1 <- lmer(richness ~ GrowthStage*Fungicide + (1|Rep), data = bull.rich.sub3)
summary(lme1)
```

- Here, Rep used as blocking factor 
- Rep is random effect.
- Results don't change drastically but we get better ability to determine the effect using mixed effect model.