---
title: "Coding challenge 1"
author: "Mamata K C"
date: "2025-02-06"
output: pdf_document
---

Introduction to R – 25 pts

This assignment is intended to introduce you to R, RStudio, and integration with GitHub, as well as help you practice some R coding. You may work in pairs. When submitting the assignment on Canvas, please put both names on the assignment. 


1.	5 pts. Explain the following things about R and Rstudio: 

o	Where do you type code? 

We type code in the top left portion of the R studio (script editor). We can also type code in the console.

o	Where is the output of the code? 

The output will appear in the console (bottom left pane of R studio)

o	Where do plots show up? 

Plots will show up in the bottom right pane of R studio in the plots tab.

o	What is a global environment? 

A global environment lies in the top right pane of the R studio. It stores all the variables, data frames, and objects that we create in R and we can view it in the environment tab. 

o	How do you get help for a function through R Studio? 

To get help for a function through R Studio, we use help(name of function) or use help tab in the bottom right pane of R Studio.

o	What is an R package? 

R package consists of many functions from different places and the R package can be downloaded from repositories. The default repository is CRAN where most of the R packages are available and can be downloaded using code install.packages(name of package).
 
o	What is a function? 

A function is a code to perform certain tasks in R.

o	How do you find the installed and loaded packages? 

To find installed and loaded packages you can go to the tab packages on the bottom left pane of R studio. If a certain package exists there and is checked then it is installed and loaded and if the package exists but is unchecked that means the package is installed but not loaded.

o	What is a working directory, and how do you find it? 

A working directory is a place or folder where you have been working and from where your files can be read and will be saved.
To find a working directory, we use the function getwd() 

o	What is a relative file path and how is it different than an absolute file path? 

Relative file path is the path of the file from your working directory whereas the absolute file path is the complete file path from root.

2.	2 pts. Explain the steps to start a new R project connected to a GitHub repository.

Go to file > click on New Project > Version Control > Git > copy-paste your GitHub repository URL and name the project and browse and set the correct folder > create

3.	3 pts. Please explain the different data classes and how they are similar and different from each other. 

o	Vector: a one-dimensional collection of elements and all the elements should be of the same data type either numeric, character, or logical.

o	Dataframe: a two-dimensional collection of elements and different columns could have a different data type however within the column the datatype should be the same.

o	Matrix: a two-dimensional collection of elements but all the elements should be of the same data type.




4.	10 pts. Complete all tasks in an R script and push it to GitHub.

o	Create a vector named 'z' with the values 1 to 200
o	Print the mean and standard deviation of z on the console
o	Create a logical vector named zlog that is 'TRUE' for z values greater than 30 and 'FALSE' otherwise.
o	Make a dataframe with z and zlog as columns. Name the dataframe zdf
o	Change the column names in your new dataframe to equal “zvec” and “zlogic”
o	Make a new column in your dataframe equal to zvec squared (i.e., z2). Call the new column zsquared. 
o	Subset the dataframe with and without the subset() function to only include values of zsquared greater than 10 and less than 100 
o	Subset the zdf dataframe to only include the values on row 26
o	Subset the zdf dataframe to only include the values in the column zsquared in the 180th row.
o	Annotate your code, commit the changes and push it to your GitHub

```{r}
#To create a vector named 'z' with the values 1 to 200.
z<-1:200  
z

#To print the mean and standard deviation of z
mean(z)
sd(z)

#To create a logical vector named zlog that is 'TRUE' for z values and 'FALSE' otherwise.
zlog<-z>30
zlog

#To make a dataframe with z and zlog as columns.Name the dataframe zdf
zdf<-data.frame(z,zlog)
zdf

#To change the column names in your new dataframe to equal "zvec" and "zlogic"
colnames(zdf)<-c("zvec","zlogic")
zdf

#To make a new column names in your dataframe equal to zvec squared (i.e, z^2). Call the new column zsquared.
zdf$zsquared<-(zdf$zvec^2)
zdf

#To subset the dataframe with the subset() function to only include values of zsquared greater than 10 and less than 100 
subset(zdf,zsquared>10 & zsquared<100)

#To subset the dataframe without the subset() function to only include values of zsquared greater than 10 and less than 100
zdf[zdf$zsquared>10 & zdf$zsquared<100,]


#To subset the zdf dataframe to only include the values on row 26
subset(zdf[26,])

#To subset the zdf dataframe to only include the values in the column zsquared in the 180th row.
zdf$zsquared[zdf$zvec==180]
```

Link to my Github repository: https://github.com/mamata2423/ReproducibilityClass/blob/main/CodingChallenge1.R




5.	5 pts. Download the Tips.csv file from Canvas. Use the read.csv() function to read the data into R so that the missing values are properly coded. **Note the missing values are reported in the data as a period (i.e., “.”). How do you know the data were read correctly? 


```{r}
read.csv("/Users/Mamata/Downloads/TipsR.csv",na.strings=".")
```

I used na.strings option to tell R that missing values in my data are denoted as "." and when I load my data, NA appears instead of "." in the missing value, we know the data is read correctly.