---
title: "Data wrangling notes"
author: "Mamata K C"
date: "2025-03-20"
output: pdf_document
toc: TRUE
---


# Data wrangling
- It is essentially idea of manipulating data for handling the large amount of messy data
- tidyverse package can be installed for data wrangling. It consists of 8 different packages.

```{r}
microbiome.data <- read.csv("Bull_richness.csv") #loading data
str(microbiome.data)
```


## Selecting certain columns


```{r}
library(tidyverse)
microbiome.data1 <- select(microbiome.data, SampleID, Crop, Compartment:Fungicide, richness) #using select function to select certain columns from the original dataset and ":" selects all the columns Compartment through Fungicide as they are in order
```


## Subsetting or filtering data 
- selecting certain rows

```{r}
head(filter(microbiome.data1, Treatment == "Conv.")) #selecting the rows only with conventional treatment & head is displaying first 6 rows
head(filter(microbiome.data1, Treatment == "Conv." & Fungicide == "C")) #selecting the rows only with conventional treatment and fungicide C
head(filter(microbiome.data1, Sample == "A" | Sample == "B")) #selectiong rows with sample A or sample B
```
## Creating a new column

```{r}
microbiome.data1$logRich <- log(microbiome.data1$richness) #previous method 

#using mutate function to create new columns
head(mutate(microbiome.data1, logRich = log(richness))) #creating new column called logRich
head(mutate(microbiome.data1, Crop_Treatment = paste(Crop, Treatment))) #creating new column combining Crop and Treatment
```



## The pipe
- allows us to combine the output from one function into the input of another function

```{r}
microbiome.data %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% #selecting columns 
  filter(Treatment == "Conv.") %>% #including only conventional treatment
  mutate(logRich = log(richness)) %>% #creating new column logRich
  head()
```


## Summarize data

```{r}
microbiome.data %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% #selecting columns 
  filter(Treatment == "Conv.") %>% #including only conventional treatment
  mutate(logRich = log(richness)) %>% #creating new column logRich
  summarise(Mean.rich=mean(logRich)) #calculating overall mean 
```
```{r}
# connect multiple summary statistics

microbiome.data %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))

```

## Group by function
 
```{r}
microbiome.data %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```
 
## Connecting to plotting
 
```{r}
microbiome.data %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot
  geom_bar(stat="identity") +
  geom_errorbar( aes(x=Fungicide, ymin=Mean.rich-std.err, ymax=Mean.rich+std.err), width=0.4) +
  theme_minimal() +
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```
 
## Joining

- allows us to combine multiple datasets based on common set of variables.
- it is important because sometimes we need to split our data to run different functions and again need to combine them to form a metadata.

```{r}
# selecting just the richness and sample ID
richness <- microbiome.data %>%
  select(SampleID, richness)

# selecting columns that don't include the richness
metadata <- microbiome.data %>% 
  select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(metadata)

head(left_join(metadata, richness, by = "SampleID")) # adding the richness data to the metadata based on on the common column of sampleID
```

## Pivoting
```{r}
microbiome.data %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) # calculate the difference between the means. 
```


```{r}
microbiome.data %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) %>%  # calculate the difference between the means. 
  ggplot(aes(x = Treatment, y = diff.fungicide)) + # Plot it 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness") 
```

