---
title: "Coding challenge 2"
author: "Mamata K C"
date: "2025-02-20"
output: pdf_document
---



Advanced Data Visualization– 25 pts



PLEASE READ THIS BEFORE CONTINUING



This assignment is designed to help you visualize data in R and practice your R coding skills. You may collaborate with a partner to enhance your learning experience. Please ensure the following:



•	Collaboration: If you worked with a partner, include their name.

•	Submission: Only one person should submit the assignment.

•	Answers: Provide written answers to all questions in the current document or a rendered markdown file.

•	Code Submission: Submit your R code in an .R or .Rmd file by sharing a link to your GitHub repository.



Partner 1:		Mamata K C					






Data description



Here, we will analyze mycotoxin data collected from wheat heads grown in the greenhouse inoculated with Fusarium graminearum (Fg) with or without fungal endophytes to protect against Fg. The experiment was performed by inoculating the heads with Fg (positive control), Fg + 1 of 3 endophytes (Fg + 40, Fg + 70, or Fg + 37), and the non-treated control (NTC). Therefore, the experiment was five treatments with 6-10 technical replicates and three experimental replicates. The authors quantified the parts per million (ppm) of mycotoxins Deoxynivalenol (DON) and the variant 15ADON from the wheat grain, and this is your primary response variable. In other words, we are interested in knowing if the fungal endophytes could reduce the concentration of DON on wheat heads. This was done across two wheat cultivars, a winter wheat variety, Ambassador, and a spring wheat variety, Wheaton. These data are published here: 



•	Noel, Z.A., Roze, L.V., Breunig, M., Trail, F. 2022. Endophytic fungi as promising biocontrol agent to protect wheat from Fusarium graminearum head blight. Plant Disease. https://doi.org/10.1094/PDIS-06-21-1253-RE

•	The following github repository is associated with this paper: https://github.com/noelzach/EndophyteBiocontrol




1.	5pts. Using ggplot, create a boxplot of DON by Treatment so that the plot looks like the image below.

a.	Jitter points over the boxplot and fill the points and boxplots Cultivar with two colors from the cbbPallete introduced last week. 

b.	Change the transparency of the jittered points to 0.6. 

c.	The y-axis should be labeled "DON (ppm)", and the x-axis should be left blank. 

d.	The plot should use a classic theme

e.	The plot should also be faceted by Cultivar


```{r}
datum <- read.csv ("MycotoxinData.csv" , na.strings = "na") #loading data in R
str(datum)  #viewing the structure of the data
library(ggplot2) #loading library ggplot2
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") #loading color blind palette

boxplot <- ggplot(data = datum, aes(x = Treatment, y = DON, fill=Cultivar)) + #defining aesthetics x as Treatment and y as DON and fill by cultivar
  geom_boxplot(position = position_dodge(), outlier.shape = NA) +  #creating box plot without overlap and removing outlier
  geom_point(position = position_jitterdodge(dodge.width = 0.8), aes(fill=Cultivar),alpha = 0.6, shape=21) + #filling the box plot with data points without overlap and setting transparency of the data points as 0.6 and shape to 21
  scale_fill_manual(values = c(cbbPalette[[4]], cbbPalette[[7]])) + #setting fills of boxplot and data points manually
  xlab("") + #labelling x axis
  ylab("DON (ppm)") + #labelling y axis
  theme_classic() + #setting theme classic to make plain white background
  facet_wrap(~Cultivar) #separating the plots according to cultivars
boxplot
```



2.	4pts. Change the factor order level so that the treatment “NTC” is first, followed by “Fg”, “Fg + 37”, “Fg + 40”, and “Fg + 70. 

```{r}
datum$Treatment <- factor(datum$Treatment, levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70")) #changing the order of the treatments as mentioned
boxplot1 <- ggplot(data = datum, aes(x = Treatment, y = DON, fill = Cultivar)) +  #defining aesthetics x as Treatment and y as DON
  geom_boxplot(position = position_dodge(), outlier.shape = NA) + #creating box plot without overlap and removing outliers
  geom_point(position = position_jitterdodge(dodge.width = 0.8), aes(fill=Cultivar),alpha = 0.6, shape=21) + #filling the box plot with data points without overlap and setting transparency of the data points as 0.6 and shape to 21
  scale_fill_manual(values = c(cbbPalette[[4]], cbbPalette[[7]])) + #setting fills of boxplot and data points manually
  xlab("") + #labelling x axis
  ylab("DON (ppm)") + #labelling y axis
  theme_classic() + #setting theme classic to make plain white background
  facet_wrap(~Cultivar) #separating the plots according to cultivars
boxplot1
```




3.	5pts. Change the y-variable to plot X15ADON and MassperSeed_mg. The y-axis label should now be “15ADON” and “Seed Mass (mg)”. Save plots made in questions 1 and 3 into three separate R objects.

```{r}

boxplot_15ADON <- ggplot(data = datum, aes(x = Treatment, y = X15ADON, fill = Cultivar)) + #defining aesthetics x as Treatment and y as DON
  geom_boxplot(position = position_dodge(), outlier.shape = NA) + #creating box plot without overlap and removing outliers
  geom_point(position = position_jitterdodge(dodge.width = 0.8), aes(fill=Cultivar),alpha = 0.6, shape=21) + #filling the box plot with data points without overlap and setting transparency of the data points as 0.6 and shape to 21
  scale_fill_manual(values = c(cbbPalette[[4]], cbbPalette[[7]])) + #setting fills of boxplot and data points manually
  xlab("") + #labelling x axis
  ylab("15ADON") + #labelling y axis
  theme_classic() + #setting theme classic to make plain white background
  facet_wrap(~Cultivar) #separating the plots according to cultivars
boxplot_15ADON

boxplot_massperseed <- ggplot(data = datum, aes(x = Treatment, y = MassperSeed_mg, fill = Cultivar)) + #defining aesthetics x as Treatment and y as DON
  geom_boxplot(position = position_dodge(), outlier.shape = NA) + #creating box plot without overlap and removing outliers
  geom_point(position = position_jitterdodge(dodge.width = 0.8), aes(fill=Cultivar),alpha = 0.6, shape=21) + #filling the box plot with data points without overlap and setting transparency of the data points as 0.6 and shape to 21
  scale_fill_manual(values = c(cbbPalette[[4]], cbbPalette[[7]])) + #setting fills of boxplot and data points manually
  xlab("") + #labelling x axis
  ylab("Seed Mass (mg)") + #labelling y axis
  theme_classic() + #setting theme classic to make plain white background
  facet_wrap(~Cultivar) #separating the plots according to cultivars
boxplot_massperseed

```


4.	5pts. Use ggarrange function to combine all three figures into one with three columns and one row. Set the labels for the subplots as A, B and C. Set the option common.legend = T within ggarage function. What did the common.legend option do?

a.	HINT: I didn’t specifically cover this in the tutorial, but you can go to the help page for the ggarange function to figure out what the common.legend option does and how to control it. 

```{r}
library(ggpubr)
figure <- ggarrange( 
  boxplot1, #plotting boxplot1
  boxplot_15ADON, #plotting boxplot_15ADON
  boxplot_massperseed, #plotting boxplot_massperseed
  labels = "AUTO", #setting labels as A, B and C
  nrow = 1, #combining figures with single row
  ncol = 3, #combining figures with 3 columns
  common.legend = TRUE, #combining legends for all three figures as one
  legend = TRUE #displaying legend
)
figure
```

 The common.legend option within ggarrange function let us combine the legends and display only one legend for multiple plots that have been combined using ggarrange function. The option can be controlled by setting it to either TRUE or FALSE. If we set the option as TRUE R will give one legend across multiple plots and if we set it as FALSE, R will give separate legends for separate plots.
 
 

5.	5pts. Use geom_pwc() to add t.test pairwise comparisons to the three plots made above. Save each plot as a new R object, and combine them again with ggarange as you did in question 4. 

```{r}
boxplot1_pwc <- boxplot1 + 
  geom_pwc(aes(group = Treatment), method = "t.test", label = "p.adj.format") #pair-wise comparisons between treatments
boxplot1_pwc

boxplot_15ADON_pwc <- boxplot_15ADON +
  geom_pwc(aes(group = Treatment), method = "t.test", label = "p.adj.format")  #pair-wise comparisons between treatments
 boxplot_15ADON_pwc

 boxplot_massperseed_pwc <- boxplot_massperseed +
  geom_pwc(aes(group = Treatment), method = "t.test", label = "p.adj.format") #pair-wise comparisons between treatments
 boxplot_massperseed_pwc
 
 figure_pwc <- ggarrange( 
  boxplot1_pwc, #plotting boxplot1_pwc
  boxplot_15ADON_pwc, #plotting boxplot_15ADON_pwc
  boxplot_massperseed_pwc, #plotting boxplot_massperseed_pwc
  labels = "AUTO", #setting labels as A, B and C
  nrow = 1, #combining figures with single row
  ncol = 3, #combining figures with 3 columns
  common.legend = TRUE, #combining legends for all three figures as one
  legend = TRUE #displaying legend
)
figure_pwc
 
```


6.	1 pt. Annotate your code, push it to GitHub. Provide me a link to your GitHub


https://github.com/mamata2423/ReproducibilityClass/blob/main/Advanced%20Visualization.Rmd


